---
title: "EF_WAE_ZM_Model1"
author: "Holly Kundel"
date: "11/15/2021"
output: html_document
---

This model is run using `ZM_LAKE` as zebra mussel invasion variable
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

Load in Required Packages
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(arm)
library(doBy)
library(sf)
library(spData)
library(RANN)
library(ggplot2)
library(MASS)
```

Load in Data
```{r, message=FALSE}
EF_WAE_ZM_data <- read_csv("EF_DATA_11.22.21.csv")

```
# Nearest Neighbor Code

- Assigning uninvaded lakes a "pseudo-invasion year" for BACI set up

Data setup
```{r}
# Format dates
dat <- EF_WAE_ZM_data %>% 
  mutate(SAMPLING_START_DATE = mdy(SAMPLING_START_DATE),
         day = day(SAMPLING_START_DATE),
         month = month(SAMPLING_START_DATE),
         year = year(SAMPLING_START_DATE),
  )
```

Create Spatial sf Object
```{r}
# Create spatial sf object (4326 = WGS84)
dat.geo <- st_as_sf(dat, coords = c("LAKE_CENTER_LONG_DD5", "LAKE_CENTER_LAT_DD5"), crs = 4326)
```

Separate Invaded Lakes and Uninvaded Lakes

- Using column `ZM_LAKE` for this
- Only considered invaded if at least one EF survey was done the year of invasion or later
- No invasions after 2019 count (because no data post 2019)
- "0" indicates a lake is uninvaded (based on above criteria)
- "1" indicates a lake is invaded (based on above criteria)

```{r}
dat.geo.uninvaded <- dat.geo %>% 
  filter(ZM_LAKE == 0)
dim(dat.geo.uninvaded) #1116 surveys

dat.geo.invaded <- dat.geo %>% 
  filter(ZM_LAKE == 1)
dim(dat.geo.invaded) #322 surveys, adds up to correct total of 1438 surveys total
```
Get coordinate matrices
```{r}
uninvaded_coords <- do.call(rbind, st_geometry(dat.geo.uninvaded))
invaded_coords <- do.call(rbind, st_geometry(dat.geo.invaded))
invaded_coords <- cbind(invaded_coords, 1:nrow(invaded_coords)) # Index for identifying nearest lake

```

Find nearest neighbor
```{r}
closest <- nn2(invaded_coords[,1:2], uninvaded_coords,  k = 1, searchtype = "standard")
str(closest)
head(closest)
closest <- sapply(closest, cbind) 
```

Combine uninvaded data with nearest neighbor assignments
```{r}
dat.geo.uninvaded <- cbind(dat.geo.uninvaded, closest)
str(dat.geo.uninvaded)
dim(dat.geo.uninvaded) 
```
Give invaded lakes an identifier and select columns
```{r}
dat.geo.invaded$nn.idx <- 1:nrow(dat.geo.invaded)
dat.geo.invaded <- dat.geo.invaded %>% 
  dplyr::select(YEAR_INFESTED, nn.idx)
str(dat.geo.invaded)
head(dat.geo.invaded)

str(dat.geo.uninvaded)
str(dat.geo.invaded)
```
Merge uninvaded with invaded `YEAR_INFESTED` for use as "after" year in BACI (Before After Control Impact)
```{r}
dat.geo.invaded_merge <- st_set_geometry(dat.geo.invaded, NULL)

dat.geo.uninvaded.final <- dat.geo.uninvaded %>%
    left_join(dat.geo.invaded_merge, by = c("nn.idx"))
str(dat.geo.uninvaded.final)
summary(dat.geo.uninvaded.final)
# Bring back into full dataset
# Convert to dataframe
dat.geo.uninvaded.final.df <- st_set_geometry(dat.geo.uninvaded.final, NULL)
class(dat.geo.uninvaded.final.df)
str(dat.geo.uninvaded.final.df)

# Join dat.geo.uninvaded.final.df - want YEAR_INFESTED.y - with original data, dat
# Grab columns of interest
dat.geo.uninvaded.final.df <- dat.geo.uninvaded.final.df %>% 
  dplyr::select(SURVEY_ID, YEAR_INFESTED.y)
dim(dat.geo.uninvaded.final.df)
head(dat.geo.uninvaded.final.df)
dim(dat)

# Final merge
final.dat3 <- dat %>%
    left_join(dat.geo.uninvaded.final.df, by = c("SURVEY_ID"))
dim(final.dat3)
head(final.dat3)
```

Export data with nearest neighbor
```{r}
write.csv(final.dat3, "NN_EF_WAE_ZM_Model_3_11.15.21.csv")
```

Test Nearest Neighbor Data
```{r}
# test
test1 <- final.dat %>%
   filter(ZM_LAKE == 0)
 head(test1)

# Export final data

# Check nearest neighbors
# Grab state boundaries from spData
us_states2 <- st_transform(us_states, crs = 4326)
# Rename column
colnames(us_states2)[2] <- "State"
selectStates <- c("Minnesota")
# Subset data for plotting
us_state_select <- us_states2[us_states2$State %in% selectStates, ]

# Subset for visualization
unique(dat.geo.uninvaded$nn.idx)
invaded.plot <- dat.geo.invaded %>% 
  filter(nn.idx == 307 | nn.idx == 309 |nn.idx == 45)

uninvaded.plot <- dat.geo.uninvaded %>% 
  filter(nn.idx == 307 | nn.idx == 309 |nn.idx == 45)

ggplot() + geom_sf(data = us_state_select, color = "gray30", lwd=1, fill="white") +
  geom_sf(data = invaded.plot, size=3.0, aes(color=as.factor(nn.idx))) +
  geom_point() +
  geom_sf(data = uninvaded.plot,size=0.3, aes(color=as.factor(nn.idx))) +
  geom_sf(data = dat.geo.invaded,size=0.3, color='black') +
  labs(title="", y="Latitude", x="Longitude", color = "Lakes") 
ggsave("NN_ZM_map.png", height=5, width=5, units="in")
```

# Prepare to run model

Load more libraries
```{r}
# if you need to install stan run: "install.packages("rstan", repos = "https://cloud.r-project.org/", dependencies = TRUE)"
library(StanHeaders)
library(rstan)
library(bayesplot)
library(lme4)
library(ggmcmc) # Use ggmcmc package for assessing convergence
library(gridExtra)
library(ggthemes)
library(coda)
library(forcats)
```

Load in Data Again 
- could also use `final.dat` from above
```{r}
dat <- read.csv('NN_EF_WAE_ZM_Model1_11.15.21.csv')
dim(dat)
head(dat)

# Ensure that dates are appropriately formatted
dat <- dat %>% 
  mutate(SAMPLING_START_DATE = ymd(SAMPLING_START_DATE),
         day = day(SAMPLING_START_DATE),
         month = month(SAMPLING_START_DATE),
         year = year(SAMPLING_START_DATE))
```

Create BACI design variables
```{r}
# "Treatment" is lake-level invasion status, ZM (1 = ZM; 0 otherwise)
# Before/After time variable ("post" = 1 if sampled after invasion; 0 if before)
# Below, "year" is year of sampling and time is year infested
dat <- dat %>% 
  mutate(time = ifelse(!is.na(YEAR_INFESTED), YEAR_INFESTED, YEAR_INFESTED.y),
         post = ifelse(year > time, 1, 0),
         treatment = ZM_LAKE # USING ZM_LAKE_HERE
         )%>%
  mutate(INVASION_STATUS_NEW = case_when(ZM_LAKE == 0 ~ "Uninvaded",
                                         ZM_LAKE == 1 & post == 0 ~ "PRE_ZM",
                                         ZM_LAKE == 1 & post == 1 ~ "POST_ZM"))

head(dat)
summary(dat)

#check data
dat %>%
  group_by(INVASION_STATUS_NEW) %>%  
  count(post, ZM_LAKE) %>%
  mutate(prop = prop.table(n))
```

Prep for model
```{r}
# Create lake ID
dat$lake.id <- as.numeric(as.factor(as.numeric(as.factor(dat$DOW_KITTLE))))
# Create year ID
dat$year.id <- as.numeric(as.factor(as.numeric(as.factor(dat$year))))
nyear <- length(unique(dat$year.id))

# Sort data by site
dat <- dat %>% 
  arrange(lake.id)

# Effort offset term (maybe multiply eff_seconds by the number_of_netters?)
dat <- dat %>% 
  mutate(EFF_HRS = EF_SECONDS/3600,
         offset = log(EFF_HRS * NUMBER_OF_NETTERS))
```

Lake-level predictors
```{r}
# Prepare lake-level predictors
lake_area <- summaryBy(AREA_HECTARES ~ lake.id, data=dat, FUN=mean)
gdd <- summaryBy(MEDIAN_93_19_GDD ~ lake.id, data=dat, FUN=mean)
secchi <- summaryBy(ANNUAL_MED_SECCHI_M ~ lake.id, data=dat, FUN=mean)

# Transform as needed and scale
lake_area2 <- as.numeric(scale(log(lake_area$AREA_HECTARES.mean)))
gdd2 <- as.numeric(scale(gdd$MEDIAN_93_19_GDD.mean))
secchi2 <- as.numeric(scale(secchi$ANNUAL_MED_SECCHI_M.mean))
```

Observation-level predictors
```{r}
# Observation-level predictors, transform as needed and scale
annual_gdd <- as.numeric(scale(dat$ANNUAL_GDD_5C)) # GDD up to day of sampling
annual_secchi <- as.numeric(scale(dat$ANNUAL_MED_SECCHI_M))
post <- dat$post
x.treatment <- dat$treatment
stocked <- dat$STOCKED
```

Quick test with glmer
```{r}
m0 <-glmer(TOTAL_CATCH ~ 1 + (1 | lake.id), data = dat, family = poisson(link = "log"), nAGQ = 100)
summary(m0)
# # 
m1 <- glmer(TOTAL_CATCH ~ post + treatment + post*treatment + STOCKED + (1|lake.id), nAGQ = 100,
             data=dat, family = poisson(link="log"))
summary(m1)
```
Check that data is in correct form
```{r}
n_post <- max(post)
n_trt <- max(x.treatment)
n_stocked <- max(stocked)
n_lake <- max(dat$lake.id)
n_year <- max(dat$year.id)
```

Load data; x = x; (dim(x)[2] + dim(x_obs)[2] + 1)
```{r}
data <- list(y = dat$TOTAL_CATCH, lake = dat$lake.id, N = dim(dat)[1], 
             n_lake = n_lake, offset=dat$offset, 
             year = dat$year.id, n_year = n_year, n_post = n_post, n_trt = n_trt, n_stocked = n_stocked,
             post = post, trt = x.treatment, stocked = stocked, survey_gdd = annual_gdd, survey_secchi = annual_secchi,
             lake_area = lake_area2, gdd = gdd2, secchi = secchi2)
```


```{r}
out <- stan(file='pois_model_zip3.stan', data=data,
            iter=50000, warmup=25000, chains=3, thin=1, cores=3, refresh = 1)
print(out)


# Save all the stan output
saveRDS(out, file="model_out_zip3.rds")
out <- readRDS(file='model_out_zip3.rds')
```



